\documentclass[a4paper,12pt]{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{braket}
\usepackage[cm]{fullpage}
\usepackage{fancyhdr}
\usepackage[compact]{titlesec}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{multirow}
\pagestyle{fancy}
\fancyhead{}
\fancyfoot{}
\fancyhead[RO,RE]{--Currell Berry--}
\fancyfoot[C]{-- Updated 9-6-2013 --}
\fancyfoot[RO, LE] {\thepage}
\newcounter{ctEqn}
\setcounter{ctEqn}{0}

%macro below: first argument: caption.  second argument: equation
\newcommand{\eqentry}[2]{
\setlength{\abovedisplayskip}{0pt}%
\setlength{\belowdisplayskip}{0pt}%
\addtocounter{ctEqn}{1}%
\begin{flalign*}
(\arabic{ctEqn}) \textbf{ #1} &&  #2%
\end{flalign*}
}

%adds some text on below an overflowing equation without starting a new
%entry of its own
\newcommand{\eqadd}[1]{
\begin{flalign*}
&&  #1
\end{flalign*}
}

\newcommand{\eqadj}[1]{\hfill{#1}}

%arguments: caption and text body
\newcommand{\textentry}[2]{
\addtocounter{ctEqn}{1}%

\noindent (\arabic{ctEqn}) \textbf{#1} #2
}

%argument: header text.  resets my counter, calls subsection
\newcommand{\msubsection}[1]{
\setcounter{ctEqn}{0}%
\subsection*{#1}
}

\newcommand{\textadd}[1]{
\hfill #1%
}

\begin{document}
\section*{\center GT Math 2403 T2 Equations Sheet}

{\setlength{\belowdisplayskip}{0pt}
\msubsection{Chapter 2}
\vspace{-4 pt}
\eqentry{definition of hbar}{\hbar = \frac{h}{2\pi} \mbox{, where h is planck's constant}}
\eqentry{uncertainty}{\sqrt{\langle (S_z)^{2} \rangle - \langle S_z \rangle ^{2}} = \Delta S_z \mbox{, where $\Delta S_z$ is uncertainty}}
\eqentry{Completeness Relation}{ \sum_{i}{}{\ket{u_i}\bra{u_i}} = 1}
switch $N_x$ and $M_y$ to test if integrating factor depends only on y.
%-----------------------Chapter 3----------------------------
\msubsection{Chapter 3: First Order Systems}
\textentry{FOLS Test for Unique Solution}{If the coefficient functions of a general linear system are continuous on an open interval I, then there exists a unique solution to the system which exists throughout the interval for any given initial values.}
\textentry{Wronskian}{If the Wronskian of two solutions is not zero, they form a fundamental set.}
\eqentry{Real and Different Eigenvalues}{y(t) = c_1e^{\lambda_1t}\boldsymbol{v_1} + c_2e^{\lambda_2t}\boldsymbol{v_2}} 
\textentry{Complex Eigenvalues}{Use Euler's formula to simplify the solution into its real part $u(t)$ and its complex part  $w(t)$.  The general solution then is $x = c_1u(t) + c_2w(t)$}
\textentry{Repeated Eigenvalues, One Eigenvector}{Find $w$ satisfying $(A - \lambda I)w = v$.}%
An additional solution is then $e^{\lambda t}(w+tv)$
%---------------------Chapter 4------------------------
\msubsection{Chapter 4: Second Order LDEs}
Find roots of characteristic polynomial.  Then use below formulas to complete the solution.
\eqentry{Real and Distinct}{c_1e^{\lambda_1t} + c_2e^{\lambda_2t}}
\eqentry{Complex}{\mbox{ Roots of form } \lambda \pm \mu i \mbox{ then real part of solution is } c_1e^{\lambda t}\cos{\mu t} + c_2e^{\lambda t} \sin{\mu t}}
\eqentry{Repeated}{c_1e^{\lambda_1t} + c_2te^{\lambda_2t}}%
\eqentry{Spring Equation}{my''(t) + \gamma y'(t) + ky(t) = F(t)}
\textentry{Damping}{$c = \gamma^2 - 4km$.  If $c<0$ then underdamped.  If $c=0$ then critically damped.  If $c>0$ then overdamped}
\textentry{Undetermined Coefficients}{}
\begin{table}[htbp]
\centering
The particular solution of $ay''+ by' + cy=g_i(t)$...\\
\begin{tabularx}{\textwidth}{>{\centering\arraybackslash}X >{\centering\arraybackslash}X }
\toprule
 $g_i(t)$ & $Y_i(t) $\\
\midrule
  $ P_n(t) = a_0t^n + a_1t^{n-1}+ ... + a_n $ & $ t^s(A_0t^n + A_1t^(n-1)+ ... + A_n $ \\
  $ P_n(t)e^{\alpha t} $ & $t^s(A_0t^n + A_1t^{n-1} + ... + A_n)e^{\alpha t}$ \\
  $
P_n(t)e^{\alpha t} \left\{
     \begin{array}{lr}
       \sin{\beta t}\\
       \cos{\beta t}\\
     \end{array}
   \right.
$ & $ t^s[(A_0t^n + A_1t^{n-1} + ... + A_n)e^{\alpha t}\cos{\beta t}+ (B_0t^n + B_1t^{n-1} + ... + B_n)e^{\alpha t}\sin{\beta t} ]$ \\
\bottomrule
\end{tabularx}
\end{table}
%\eqadd{\mbox{where X is } [x_1,x_2] \mbox{ \& } \{x_1,x_2\}   \mbox{ form a fundamental set of solution to } x'=P(t)x }
\textentry{Variation of Parameters 1}{Given $[x_1,x_2]$ which form a fundamental set of solutions to $x'=P(t)x$, then $X=\{x_1,x_2\}$, and a particular solution to $y''+p(t)y'+q(t)y=g(t)$ is \fbox{$X(t)\int{X^{-1}(t)g(t) dt}$}}
\textentry{Variation of Parameters 2}{Given solutions $y_1$ and $y_2$ to the corresponding homogenous equation, a particular solution to $y''+p(t)y'+q(t)y=g(t)$ is}
\eqadd{Y(t)=-y_1(t)\int{\frac{y_2(t)g(t)}{W[y_1,y_2](t)} dt} + \int{\frac{y_1(t)g(t)}{W[y_1,y_2](t)} dt}}
%---------------------Chapter 5------------------------
\msubsection{Chapter 5: Laplace Transformations}
\renewcommand{\arraystretch}{1.5}
\eqentry{Definition of Laplace Transform}{\int_0^\infty{e^{-st}f(t) dt}}

\begin{tabularx}{\textwidth}{ X X X X }
\multicolumn{4}{ c }{Table of Common Laplace Transformations} \\ \hline
$f(t) = \mathcal{L}^{-1}\{F(s)\}$ & $F(s) = \mathcal{L}\{f(t)\}$ & $f(t) = \mathcal{L}^{-1}\{F(s)\}$ & $F(s) = \mathcal{L}\{f(t)\}$ \\
\hline
$1$&	$\frac{1}{s}$&	$e^{ct}f(t)$&	$F(s-c)$\\
$t^n$&	$\frac{n!}{s^{n+1}}$&	$t^nf(t)$&	$(-1)^nF^{(n)}(s)$\\
$\sin (at)$&	$\frac{a}{s^2+a^2}$&	$\cos {at}$&	$\frac{s}{s^2+a^2}$\\
$\sinh {at}$&	$\frac{a}{s^2- a^2}$&	$\cosh {at}$&	$\frac{s}{s^2-a^2}$\\
$u_c(t)f(t-c)$&	$e^{-cs}F(s)$&	$\delta(t-c)$&	$e^{-cs}$\\
$\int_0^t {f(t-\tau)g(\tau) d\tau}$&	$F(s)G(s)$&	$f'(t)$&	$sF(s)-f(0)$\\

\end{tabularx}
%---------------------Chapter 7 ------------------------
\msubsection{Chapter 7: Nonlinear Differential Equations and Stability }
\textentry{Condition for Linear Approximation}{Given general nonlinear system $x'=F(x,y)$, $y'=G(x,y)$ (A).  The system is almost linear in the neighborhood of any isolated critical point $(x_0,y_0)$ whenever the functions F and G have continuous partial derivatives up to order 2.}
\textentry{Almost Linear Approximation}{Find the Jacobian matrix of the system.  Plug in the x and y values of each critical point for a linear approximation of the system near that critical point.}
\end{document}
